---
title: "R2_dataset_kaplan_meier_curve"
author: "JunhuiLi"
date: "2023-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Kaplan meier curve for 3 dataset

We made kaplan meier curve for overall time - status ~ C1GALT1 and eventfree - status ~  C1GALT1 based 3 dataset(GSE17679, GSE63157_core_exon and GSE63157_core_transcript). 

we compared if there is a difference in eventfree or overall between two groups with C1GALT1 expression high and low. The cutoff methods grouping samples into high and low includes scan, median, average, first_quartile, last_quartile, last_vs_firt_quartile and curtain.

Note that the p value from survfit in R is different from R2 platform.

```{r}
library(survminer)
library(survival)
library(ggplot2)
library(patchwork)
```

## For dataset gse17679

prepare the input raw dataset
```{r}
gse17679_raw <- read.table("result/gse17679/gse17679.txt",comment.char = "")
gse17679_raw[1:15,1:5]

gse17679_C1GALT1 <- gse17679_raw[c(1:10,which(gse17679_raw[,1] %in% c("C1GALT1","SMO","GLI1"))),]
gse17679 <- as.data.frame(t(gse17679_C1GALT1))
gse17679[1:5,]
colnames(gse17679) <- c("hugo","age","efs","id","overall","sample_series_id","set","sex","state","status","C1GALT1","GLI1","SMO")

dat <- gse17679[-c(1:2),]
dat[1:5,]
dim(dat)

dat_remove_nd <- dat[!dat$efs %in% "nd",]

dat_remove_nd$age <- as.numeric(dat_remove_nd$age)
dat_remove_nd$efs <- as.numeric(dat_remove_nd$efs)
dat_remove_nd$overall <- as.numeric(dat_remove_nd$overall)
dat_remove_nd$C1GALT1_value <- as.numeric(dat_remove_nd$C1GALT1)
str(dat_remove_nd)
dat_remove_nd_sort <- dat_remove_nd[order(dat_remove_nd$C1GALT1_value, decreasing = FALSE),]
dat_remove_nd_sort$id <- 1:nrow(dat_remove_nd_sort)
dat_remove_nd_sort$Event[dat_remove_nd_sort$status=="dead"] <- "YES"
dat_remove_nd_sort$Event[dat_remove_nd_sort$status!="dead"] <- "NO"
dat_remove_nd_sort$Event <- as.factor(dat_remove_nd_sort$Event)
dat_remove_nd_sort$status
head(dat_remove_nd_sort)
str(dat_remove_nd_sort)
```
correlation plot
```{r}
dat[1:5,]
dat_corr <- dat
dat_corr$C1GALT1 <- as.numeric(dat_corr$C1GALT1)
dat_corr$C1GALT1_log2 <- log2(dat_corr$C1GALT1)
summary(dat_corr$C1GALT1)
dat_corr$GLI1 <- as.numeric(dat_corr$GLI1)
dat_corr$GLI1_log2 <- log2(dat_corr$GLI1)
summary(dat_corr$GLI1)
dat_corr$SMO <- as.numeric(dat_corr$SMO)
dat_corr$SMO_log2 <- log2(dat_corr$SMO)
summary(dat_corr$SMO)
str(dat_corr)


pc1 <- ggplot(data=dat_corr,aes(x=C1GALT1_log2,y=GLI1_log2))+
  geom_point()+
  theme_classic2()+
  stat_cor(method = "pearson", label.y = max(dat_corr$GLI1_log2)+1)

pc2 <- ggplot(data=dat_corr,aes(x=C1GALT1_log2,y=SMO_log2))+
  geom_point()+
  theme_classic2()+
  stat_cor(method = "pearson", label.y = max(dat_corr$SMO_log2)+1)

pdf(paste0("result/gse17679/correlation/gse17679_C1GALT1_GLI1_correlation.pdf"))
    print(pc1)
dev.off()
pdf(paste0("result/gse17679/correlation/gse17679_C1GALT1_SMO_correlation.pdf"))
    print(pc2)
dev.off()
# install.packages("PerformanceAnalytics")
library(PerformanceAnalytics)
dat_corr_gene <- dat_corr[,c("C1GALT1_log2","GLI1_log2","SMO_log2")]
pdf(paste0("result/gse17679/correlation/gse17679_C1GALT1_GLI1_SMO_correlation.pdf"))
    chart.Correlation(dat_corr_gene, histogram = TRUE, method = "pearson")
dev.off()
```



go through all cutoff methods.
```{r}
## scan threshold
cutoff <- list()
cutoff["scan"]  <- list(63.1)
cutoff["median"]  <- list(81.8)
cutoff["average"]  <- list(87.1)
cutoff["first_quartile"]  <- list(54.4)
cutoff["last_quartile"]  <- list(109.9)
cutoff["first_vs_last_quartile"]  <- list(23)
cutoff["curtain"]  <- list(14)

#i=names(cutoff)[7]
for(i in names(cutoff)){
  cut_off <- cutoff[[i]]
  if(i=="first_vs_last_quartile" | i=="curtain"){
    dat_remove_nd_sort[(cut_off+1):nrow(dat_remove_nd_sort),"C1GALT1"] <- "high"
    dat_remove_nd_sort[1:cut_off,"C1GALT1"] <- "low"
    cut_off <- dat_remove_nd_sort[cut_off,"C1GALT1_value"]
  }else{
    dat_remove_nd_sort[dat_remove_nd_sort$C1GALT1_value > cut_off,"C1GALT1"] <- "high"
    dat_remove_nd_sort[dat_remove_nd_sort$C1GALT1_value <= cut_off,"C1GALT1"] <- "low"
    dat_remove_nd_sort$C1GALT1 <- as.factor(dat_remove_nd_sort$C1GALT1)
  }
    
  #Y = Surv(dat_remove_nd_sort$overall, dat_remove_nd_sort$status == "dead")
  #kmfit = survfit(Y ~ dat_remove_nd_sort$C1GALT1)
  #summary(kmfit, times = c(seq(0, 1000, by = 100)))
  
  fit_overall <- survfit(Surv(overall, status== "dead") ~ C1GALT1,
                 data = dat_remove_nd_sort)

  p1_overall <- ggsurvplot(fit_overall,
                           legend.title = "",
                           palette = c("darkred", "darkblue"),
                           data = dat_remove_nd_sort, 
                           risk.table = TRUE,
                           pval = TRUE,
                           surv.plot.height=0.85)+
    xlab("Follow up in months")+
    ylab("Overall survival probability")
  
  p2 <- ggplot(data=dat_remove_nd_sort,aes(x=id,y=C1GALT1_value,color=Event))+
    geom_point()+
    theme_classic()+
    geom_vline(xintercept = sum(dat_remove_nd_sort$C1GALT1=="low"), color="gray")+
    geom_hline(yintercept = cut_off, color="gray")+
    ylab("Expression")+
    xlab("sorted by expression")+
    theme(axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          legend.position = "top")
  
  
  #(p1$plot/p1$table)|p2
  pdf(paste0("result/gse17679/gse17679_2plots_cutoff_method_",i,".overall.pdf"))
    plot_layout(widths = c(2, 1), heights = c(1, 1))
    print(p1_overall$plot + p2 +  plot_layout(widths = c(3, 1)))
  dev.off()
  
  layout <- "
  AAAAAACC
  AAAAAACC
  AAAAAACC
  AAAAAACC
  AAAAAACC
  AAAAAACC
  AAAAAACC
  BBBBBBCC
  "
  
  pdf(paste0("result/gse17679/gse17679_3plots_cutoff_method_",i,".overall.pdf"))
    print(p1_overall$plot + p1_overall$table + p2 + 
    plot_layout(design = layout))
  dev.off()
  
  
  fit_efs <- survfit(Surv(efs, status== "dead") ~ C1GALT1,
                 data = dat_remove_nd_sort)
  
  p1_efs <- ggsurvplot(fit_efs,
                       legend.title = "",
                       data = dat_remove_nd_sort,
                       palette = c("darkred", "darkblue"),
                       risk.table = TRUE,
                       pval = TRUE,
                       surv.plot.height=0.85)+
    xlab("Follow up in months")+
    ylab("eventfree survival probability")
  
  #(p1$plot/p1$table)|p2
  pdf(paste0("result/gse17679/gse17679_2plots_cutoff_method_",i,".eventfree.pdf"))
    plot_layout(widths = c(2, 1), heights = c(1, 1))
    print(p1_efs$plot + p2 +  plot_layout(widths = c(3, 1)))
  dev.off()
  pdf(paste0("result/gse17679/gse17679_3plots_cutoff_method_",i,".eventfree.pdf"))
    print(p1_efs$plot + p1_efs$table + p2 + 
    plot_layout(design = layout))
  dev.off()
}
```
